\documentclass[DIV=14,titlepage=false]{scrreprt}

\input{preamble.tex}
\title{%
R300 Econometrics}
\author{Metrics Enjoyers}
\date{Michaelmas Term, 2023-2024}


\setuptoc{toc}{leveldown}


\begin{document}

\maketitle

\tableofcontents
\newpage
\chapter{Lecture 1: Basic Probability. Conditional expectation function.}
\section{Review of continuous distributions}

\begin{definition}[Cumulative distribution function]
  The cumulative distribution function of X is  defined as $ F_X(x) \equiv P(X \leq x) $. 
  By the Fundamental Theorem of Calculus, $ \frac{d}{dx} F_X(x) = f_X(x) $ for a continuous r.v.\ at continuity points of $ f_X $. A function $ F $ is a cdf iff:
  \begin{enumerate}
  \item $ \lim_{x \to -\infty} F(x) = 0 $ and $ \lim_{x \to \infty} F(x) = 1 $;
  \item $ F (\cdot) $ nondecreasing;
  \item $ F (\cdot) $ right-continuous; i.e., $ \forall x_0 $, $ \lim_{x \downarrow x_0} F(x) = F(x_0) $.
  \end{enumerate}
\end{definition}

\section{Conditional expectation function (CEF)}
\subsection{Conditional quantile function}
\chapter{Causal interpretation of regression. Least Squares.}
\section{Regression and causality}
\section{Estimating population regression by least squares}

\newpage
% Nicely formatted boxes
\begin{theorem}
  This is a theorem.
  \end{theorem}
  \begin{proof}[Theorem 1]
  This is a proof.
  \end{proof}
  \begin{example}
  This is an example.
  \end{example}
  \begin{explanation}
  This is an explanation.
  \end{explanation}
  \begin{claim}
  This is a claim.
  \end{claim}
  \begin{corollary}
  This is a corollary.
  \end{corollary}
  \begin{prop}
  This is a proposition.
  \end{prop}
  \begin{lemma}
  This is a lemma.
  \end{lemma}
  \begin{question}
  This is a question.
  \end{question}
  \begin{solution}
  This is a solution.
  \end{solution}
   \begin{question}
  This is another question.
  \end{question}
  \begin{solution}
  This is another solution.
  \end{solution}
  \begin{exercise}
  This is an exercise.
  \end{exercise}
  \begin{definition}[Test]
  This is a definition.
  \end{definition}
  \begin{note}
  This is a note.
  \end{note}

\end{document}
